EinsteinPy: Python for General Relativity

### 2020-01-24 authors.tex
Fix things

% \author[0000-0000-0000-0000]{The EinsteinPy Community}
% \noaffiliation

@@

\author{Ritwik Saha}
\altaffiliation{EinsteinPy Lead Developer, EinsteinPy Board Member}
\affiliation{\sceeiitmandi}

\author{Bhavya Bhatt}
\altaffiliation{EinsteinPy Deputy Lead Developer}
\affiliation{\sceeiitmandi}

@@

% Below add contributors to the EinsteinPy project

@@

@@

@@

\author[0000-0001-5558-9447]{Raphael Reyna}
\affiliation{California State Polytechnic University, Pomona}

\author[0000-0001-8302-1584]{Sashank Mishra}
\affiliation{Department of Information Technology, Indian Institute of Information Technology Allahabad}

>>>>

% \author[0000-0000-0000-0000]{The SunPy Community}
% \noaffiliation

@@

\author[0000-0002-2369-0628]{Ritwik Saha}
\altaffiliation{EinsteinPy Lead Developer, EinsteinPy Board Member}
\affiliation{\sceeiitmandi}

\author[0000-0002-2784-5666]{Bhavya Bhatt}
\altaffiliation{EinsteinPy Deputy Lead Developer}
\affiliation{\sceeiitmandi}

@@

% Below add contributors to the Sunpy project
% this list is ordered alphabetically

@@

\author[0000-0002-1928-1549]{Rishi Sharma}
\altaffiliation{EinsteinPy Board Member}
\affiliation{\sceeiitmandi}

@@

\author[0000-0003-4387-8078]{Divya Gupta}
\affiliation{\sceeiitmandi}

\author{Hrishikesh Sudam Sarode}
\affiliation{\sceeiitmandi}

@@

\author[0000-0001-5558-9447]{Raphael Reyna}
\affiliation{California State Polytechnic University, Pomona, USA}

\author[0000-0002-6696-8730]{Govind Dixit}
\affiliation{Information Technology, Indian Institute of Information Technology Lucknow, India}

\author{Ratin Kumar}
\affiliation{National Institute of Technology Kurukshetra, India}

\author[0000-0001-8302-1584]{Sashank Mishra}
\affiliation{Department of Information Technology, Indian Institute of Information Technology Allahabad, India}

\author[0000-0003-0973-4994]{Pratyush Kerhalkar}
\affiliation{Department of Electronics & Communication Engineering, Manipal Institute of Technology, India}

### 2020-01-24 paper.tex
Fix things

\title{EinsteinPy: Python for General Relativity\footnote{Released on July, 15th, 2019}}

@@

\subsection{Body} \label{subsec:Body}
The Body data type helps define the central black-hole or the heavy massed object and the objects revolving around it. A new body can be defined by some of the very basic constraints:

\begin{itemize}
\item \texttt{name}: Body Name,
\item \texttt{mass}: Mass of the body with attached astropy units 
\item \texttt{R}: Radius of the body with attached astropy units
\item \texttt{differential}: Complete position and velocity vector of the body in form of a Cartesian / Spherical / BoyerLindquist data class object from einsteinpy coordinates (optional)
\item \texttt{a}: Spin factor (a) of the body for Kerr Solutions with attached astropy units (optional)
\item \texttt{q}: Charge of the body with attached astropy units (optional)
\item \texttt{is\_attractor}: A boolean variable concerned with if the body is to be supposed as a central attractor. (optional)
\item \texttt{parent}: Parent body for the body, object of the same data class (optional)
\end{itemize}

\subsection{Geodesic} \label{subsec:Geodesic}
The Body data type helps define the central black-hole or the heavy massed object and the objects revolving around it. A new body can be defined by some of the very basic constraints:

@@

This class along with its parent class \textit{Tensor} have the following functions/properties,
\begin{itemize}
\item \texttt{order} (Property) : Returns the order of the tensor. 
\item \texttt{config} (Property) : Returns the arrangement of contravariant(upper)/covariant(lower) indices of the tensor. For example, 'ulll'.
\item \texttt{parent\_metric} (Property) : Returns the metric tensor of the space-time associated with the tensor, if present. Else returns \texttt{None}.
\item \texttt{tensor} (Function) : Returns the tensorial quantity contained within the class instance.
\item \texttt{subs} (Function) : Returns a new tensor upon substitution of current variables with new variables/values.
\item \texttt{simplify} (Function) : Returns a tensor contain simplified expressions from the current tensor.
\item \texttt{symbols} (Function) : Returns the symbols denoting each axis of the space-time.
\item \texttt{tensor\_lambdify} (Function) : Returns lambdified function of symbolic tensors. This means that the returned functions can accept numerical values and return numerical quantities.
\item \texttt{lorentz\_transform} (Function) : Returns a lorentz-transformed tensor when supplied with a transformation matrix. This function also considers the contravariance and covariance of tensors during transformation.
\end{itemize}

@@

The same equation can be written in a tensorial form as follows,

\begin{gather}
g_{\mu \nu} = \begin{bmatrix} - (1 - \frac{r_s}{r})c^2 & 0 & 0 & 0 \\ 0 & \frac{1}{(1 - \frac{r_s}{r})} & 0 & 0 \\ 0 & 0 & r^2 & 0 \\ 0 & 0 & 0 &  r^2 sin^2 \theta \end{bmatrix} \\ 
\nonumber \\
ds^2 = g_{\mu \nu} dx^\mu dx^\nu
\end{gather}

where $r_s$ is the schwarzschild radius of a given heavy body and is given by the equation,

\begin{gather}
r_s = \frac{2GM}{c^2}
\end{gather}

$G$, $M$, and $c$ are gravitational constant, mass of the body and speed of light in vacuum respectively.

Schwarzschild Radius is a limit below which the degree of space-time curvature causes the partcle to undergo irreversible gravitational collapse wherein the partcle would need a speed greater than speed of light ot escape singularity.

@@

The file \textbf{schwarzschild.py} contains the class \textit{Schwarzschild} for defining Schwarzschild geometry methods.

Here is a working code describing the motion of particle revolving around a black hole of mass approximately equal to mass of the Earth. As the trajectory starts from a mere $130m$ from the black hole, we obtain some interesting observations.

\begin{lstlisting}[language=Python, caption=Importing the required modules]
>>> import numpy as np
>>> import astropy.units as u
>>> from einsteinpy.plotting import GeodesicPlotter
>>> from einsteinpy.coordinates import SphericalDifferential
>>> from einsteinpy.bodies import Body
>>> from einsteinpy.geodesic import Geodesic
>>> from einsteinpy.metric import Schwarzschild
\end{lstlisting}

\begin{lstlisting}[language=Python, caption=Defining the attractor and the particle and calculating geodesics]
>>> Attractor = Body(name="BH", mass=6e24 * u.kg)
>>> init_conditions = SphericalDifferential(130*u.m, np.pi/2*u.rad, -np.pi/8*u.rad,
...                                         0*u.m/u.s, 0*u.rad/u.s, 1900*u.rad/u.s)
>>> Particle = Body(differential=init_conditions, parent=Attractor)
>>> geodesic = Geodesic(body=Particle, time=0 * u.s, end_lambda=0.00205, step_size=5e-8,
...                     metric=Schwarzschild)
>>> geodesic.trajectory
array([[ 0.00000000e+00,  1.20104339e+02, -4.97488462e+01, ...,
         9.45228078e+04,  2.28198245e+05,  0.00000000e+00],
       [ 3.99986427e-08,  1.20108103e+02, -4.97397110e+01, ...,
         9.36472607e+04,  2.28560869e+05, -5.80280784e-14],
       ...,
       [ 2.04962579e-03,  1.04114904e+02,  1.08255773e+01, ...,
        -1.23245830e+06,  1.80261847e+05, -9.67931265e-09],
       [ 2.05002578e-03,  1.03619005e+02,  1.08973776e+01, ...,
        -1.24706334e+06,  1.78734562e+05, -9.79337222e-09]])
>>> geodesic.trajectory.shape
(6567, 8)
\end{lstlisting}
We can see that for each timestep, we obtain eight values : $t$, $r$, $\theta$, $\phi$, $v_t$, $v_r$, $v_\theta$ and $v_\phi$ corrosponding to coordinate time, radial distance from the centre, polar angle, azimuth angle and their respective velocities w.r.t proper time $\tau$.

Further, we plot the trajectory in 2D for vizualization purposes.
\begin{lstlisting}[language=Python, caption=Plotting the trajectory]
>>> obj = GeodesicPlotter()
>>> obj.plot(geodesic)
>>> obj.show()
\end{lstlisting}

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.7]{images/perihelion_cropped.png}
	%\centerline{\includegraphics{ftcbuild2.pdf}}
	\caption{Plot obtained from the above code showing the phenomenon of perihelion advance}
	\label{fig:schwarzschild}
\end{figure}

We observe the phenomenon of perihelion advance of a particle around a heavy body due to spacetime curavture. The orbit would have been purely eliptical if plotted in accordance with Newtonian Mechanics, however, the curvature of space-time around the heavy mass shifts the orbit with each revolution. In practicality, this phenomenon is not so pronounced as no particle approaches a heavy mass this closely. For instance, Mercury advances its perihelion at a rate of $42.98\pm 0.04$ arcsec/century.

@@

The file \textbf{kerr.py} contains the class \textit{Kerr} for defining Schwarzschild geometry methods.

@@

\begin{lstlisting}[language=Python, caption=Defining the Kerr Body and the particle and calculating geodesics]
>>> a = 0.3 * u.m
>>> Attractor = Body(name="BH", mass=1.989e30 * u.kg, a=a)
>>> init_conditions = BoyerLindquistDifferential(49.95e5 * u.km, np.pi / 2 * u.rad, 
...                                              np.pi * u.rad, 0 * u.km / u.s, 
...                                              0 * u.rad / u.s, 0 * u.rad / u.s,
...                                              a)
>>> Particle = Body(differential=init_conditions, parent=Attractor)
>>> geodesic = Geodesic(body=Particle, time=0 * u.s, end_lambda=33932.90,
...                     step_size=1.2, metric=Kerr)
\end{lstlisting}

As evident from the code, API across different classes are made to be as homegenous as possible.

As a rotating black hole drags the space-time around it, a freely falling object acquires a angular velocity around the coordinate centre, i.e centre of the black hole. This effect is evident from the plot obatined from the above geodesics, in the same way as obtained one in the Schwarzschild example.

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.1]{images/blank.jpg} %replace this when the actual plot becomes available
	%\centerline{\includegraphics{ftcbuild2.pdf}}
	\caption{Plot obtained from the above geodesics showing the phenomenon of frame dragging}
	\label{fig:kerr}
\end{figure}

\subsubsection{Kerr-Newman Metric}\label{subsubsec:kerrnewman}
Kerr-Newman metric presents the most general case within single-body vacuum solutions of Einstein Field Equation. It generalizes all the quantities used to describe a black hole, as affirmed by "no hair" theoram, i.e. Charge $Q$, Angular Momentum $J$ and Mass $M$. The metric equation is given by,

@@

\subsection{COORDINATES}
\subsection{SYMBOLIC}
General Relativity required heavy mathematical formulation to derive and describe various tensors attributing to various features essential for cogent description of an arbitrary space-time. We would describe the various quantities currently supported by this module while maintaining brevity. Each quantity has its own class, even scalars, which are treated as 0\textsuperscript{th} order tensors.

@@

The defined class functions are :
\begin{itemize}
\item \texttt{change\_config} : returns a different instance of the same class with different index (contravariant or covariant) configuration. 
\item \texttt{inv} : returns inverse of the metric
\end{itemize}

Also, the library maintains an ever expanding predefined metrics for direct use. Here's an code snippet for a better dive into the specificities of the class.

\begin{lstlisting}[language=Python, caption=Importing a predefined metric]
>>> from einsteinpy.symbolic.predefined import CMetric
>>> metric = CMetric
>>> metric.tensor()
\end{lstlisting}

\begin{center}
$\displaystyle \left[\begin{matrix}- \frac{h{\left(y \right)}}{\left(x + y\right)^{2}} & 0 & 0 & 0\\0 & \frac{1}{\left(x + y\right)^{2} f{\left(x \right)}} & 0 & 0\\0 & 0 & \frac{1}{\left(x + y\right)^{2} h{\left(y \right)}} & 0\\0 & 0 & 0 & \frac{f{\left(x \right)}}{\left(x + y\right)^{2}}\end{matrix}\right]$
\end{center}

\begin{lstlisting}[language=Python, caption=Displaying the contravarint configuration of metric]
>>> metric.change_config("uu").tensor()
\end{lstlisting}

\begin{center}
 $\displaystyle \left[\begin{matrix}- \frac{\left(x + y\right)^{2}}{h{\left(y \right)}} & 0 & 0 & 0\\0 & \left(x + y\right)^{2} f{\left(x \right)} & 0 & 0\\0 & 0 & \left(x + y\right)^{2} h{\left(y \right)} & 0\\0 & 0 & 0 & \frac{\left(x + y\right)^{2}}{f{\left(x \right)}}\end{matrix}\right]$ \\
\end{center}

@@

Instance of the class can be made by directly passing a 3\textsuperscript{rd} order tensor or can be obtained using the following classmethods,
\begin{itemize}
\item \texttt{from\_metric} : Calculates christoffel symbols from a given instance of \hyperref[subsubsec:metric]{\textit{MetricTensor}} class.
\end{itemize}

The defined class function(s) are : 
\begin{itemize}
\item \texttt{change\_config} : returns a different instance of the same class with different index (contravariant or covariant) configuration. 
\end{itemize}

@@

Instance of the class can be made by directly passing a 4\textsuperscript{th} order tensor or can be obtained using the following classmethods
\begin{itemize}
\item \texttt{from\_metric} : Calculates Riemann Tensor from a given instance of \hyperref[subsubsec:metric]{\textit{MetricTensor}} class.
\item \texttt{from\_christoffels} : Calculates Riemann Tensor from a given instance of \hyperref[subsubsec:chl]{\textit{ChristoffelSymbols}} class.
\end{itemize}

The defined class function(s) are : 
\begin{itemize}
\item \texttt{change\_config} : returns a different instance of the same class with different index (contravariant or covariant) configuration. 
\end{itemize}

@@

Instance of the class can be made by directly passing a 2\textsuperscript{nd} order tensor or can be obtained using the following classmethods
\begin{itemize}
\item \texttt{from\_metric} : Calculates Ricci Tensor from a given instance of \hyperref[subsubsec:metric]{\textit{MetricTensor}} class.
\item \texttt{from\_christoffels} : Calculates Ricci Tensor from a given instance of \hyperref[subsubsec:chl]{\textit{ChristoffelSymbols}} class.
\item \texttt{from\_riemann} : Calculates Ricci Tensor from a given instance of \hyperref[subsubsec:riemann]{\textit{RiemannCurvatureTensor}} class.
\end{itemize}

The defined class function(s) are : 
\begin{itemize}
\item \texttt{change\_config} : returns a different instance of the same class with different index (contravariant or covariant) configuration. 
\end{itemize}

@@

Instance of the class can be made by directly passing a symbolic expression or can be obtained using the following classmethods
\begin{itemize}
\item \texttt{from\_metric} : Calculates Ricci Scalar from a given instance of \hyperref[subsubsec:metric]{\textit{MetricTensor}} class.
\item \texttt{from\_christoffels} : Calculates Ricci Scalar from a given instance of \hyperref[subsubsec:chl]{\textit{ChristoffelSymbols}} class.
\item \texttt{from\_riemann} : Calculates Ricci Scalar from a given instance of \hyperref[subsubsec:riemann]{\textit{RiemannCurvatureTensor}} class.
\item \texttt{from\_riccitensor} : Calculates Ricci Scalar from a given instance of \hyperref[subsubsec:riccitensor]{\textit{RicciTensor}} class.
\end{itemize}

The defined class properties are : 
\begin{itemize}
\item \texttt{expr} : returns the scalar symbolic expression contained in the class. 
\end{itemize}

@@

\subsection{HYPERSURFACE}
\subsection{PLOTTING}

@@

\subsection{ Documentation and gallery} \label{subsec:doc}

@@

We thank all the people that have made this AASTeX what it is today.  This
includes but not limited to Bob Hanisch, Chris Biemesderfer, Lee Brotzman,
Pierre Landau, Arthur Ogawa, Maxim Markevitch, Alexey Vikhlinin and Amy
Hendrickson. Also special thanks to David Hogg and Daniel Foreman-Mackey
for the new "modern" style design. Considerable help was provided via bug
reports and hacks from numerous people including Patricio Cubillos, Alex
Drlica-Wagner, Sean Lake, Michele Bannister, Peter Williams, and Jonathan
Gagne.

@@

\vspace{5mm}
\facilities{HST(STIS), Swift(XRT and UVOT), AAVSO, CTIO:1.3m,
CTIO:1.5m,CXO}

@@

\appendix

\section{Appendix information}

Appendices can be broken into separate sections just like in the main text.
The only difference is that each appendix section is indexed by a letter
(A, B, C, etc.) instead of a number.  Likewise numbered equations have
the section letter appended.  Here is an equation as an example.

\begin{equation}
I = \frac{1}{1 + d_{1}^{P (1 + d_{2} )}}
\end{equation}

Appendix tables and figures should not be numbered like equations. Instead
they should continue the sequence from the main article body.

>>>>

\title{EinsteinPy: Python for General Relativity}

@@

\subsection{Body} \label{subsec:Body}
The Body data type helps define the central black-hole or the heavy massed object and the objects revolving around it. 
% A new body can be defined by some of the very basic constraints:

% \begin{itemize}
% \item \texttt{name}: Body Name,
% \item \texttt{mass}: Mass of the body with attached astropy units 
% \item \texttt{R}: Radius of the body with attached astropy units
% \item \texttt{differential}: Complete position and velocity vector of the body in form of a Cartesian / Spherical / BoyerLindquist data class object from einsteinpy coordinates (optional)
% \item \texttt{a}: Spin factor (a) of the body for Kerr Solutions with attached astropy units (optional)
% \item \texttt{q}: Charge of the body with attached astropy units (optional)
% \item \texttt{is\_attractor}: A boolean variable concerned with if the body is to be supposed as a central attractor. (optional)
% \item \texttt{parent}: Parent body for the body, object of the same data class (optional)
% \end{itemize}

\subsection{Geodesic} \label{subsec:Geodesic}
The Body data type helps define the central black-hole or the heavy massed object and the objects revolving around it. 
% A new body can be defined by some of the very basic constraints:

@@

% This class along with its parent class \textit{Tensor} have the following functions/properties,
% \begin{itemize}
% \item \texttt{order} (Property) : Returns the order of the tensor. 
% \item \texttt{config} (Property) : Returns the arrangement of contravariant(upper)/covariant(lower) indices of the tensor. For example, 'ulll'.
% \item \texttt{parent\_metric} (Property) : Returns the metric tensor of the space-time associated with the tensor, if present. Else returns \texttt{None}.
% \item \texttt{tensor} (Function) : Returns the tensorial quantity contained within the class instance.
% \item \texttt{subs} (Function) : Returns a new tensor upon substitution of current variables with new variables/values.
% \item \texttt{simplify} (Function) : Returns a tensor contain simplified expressions from the current tensor.
% \item \texttt{symbols} (Function) : Returns the symbols denoting each axis of the space-time.
% \item \texttt{tensor\_lambdify} (Function) : Returns lambdified function of symbolic tensors. This means that the returned functions can accept numerical values and return numerical quantities.
% \item \texttt{lorentz\_transform} (Function) : Returns a lorentz-transformed tensor when supplied with a transformation matrix. This function also considers the contravariance and covariance of tensors during transformation.
% \end{itemize}

@@

%The same equation can be written in a tensorial form as follows,

%\begin{gather}
%g_{\mu \nu} = \begin{bmatrix} - (1 - \frac{r_s}{r})c^2 & 0 & 0 & 0 \\ 0 & \frac{1}{(1 - \frac{r_s}{r})} & 0 & 0 \\ 0 & 0 & r^2 & 0 \\ 0 & 0 & 0 &  r^2 sin^2 \theta \end{bmatrix} \\ 
%\nonumber \\
%ds^2 = g_{\mu \nu} dx^\mu dx^\nu
%\end{gather}

%where $r_s$ is the schwarzschild radius of a given heavy body and is given by the equation,

%\begin{gather}
%r_s = \frac{2GM}{c^2}
%\end{gather}

%$G$, $M$, and $c$ are gravitational constant, mass of the body and speed of light in vacuum respectively.

Schwarzschild Radius is a limit below which the degree of space-time curvature causes the particle to undergo irreversible gravitational collapse wherein the particle would need a speed greater than speed of light ot escape singularity.

@@

%Here is a working code describing the motion of particle revolving around a black hole of mass approximately equal to mass of the Earth. As the trajectory starts from a mere $130m$ from the black hole, we obtain some interesting observations.

% \begin{lstlisting}[language=Python, caption=Importing the required modules]
% >>> import numpy as np
% >>> import astropy.units as u
% >>> from einsteinpy.plotting import GeodesicPlotter
% >>> from einsteinpy.coordinates import SphericalDifferential
% >>> from einsteinpy.bodies import Body
% >>> from einsteinpy.geodesic import Geodesic
% \end{lstlisting}

% \begin{lstlisting}[language=Python, caption=Defining the attractor and the particle and calculating geodesics]
% >>> Attractor = Body(name="BH", mass=6e24 * u.kg)
% >>> init_conditions = SphericalDifferential(130*u.m, np.pi/2*u.rad, -np.pi/8*u.rad,
% ...                                         0*u.m/u.s, 0*u.rad/u.s, 1900*u.rad/u.s)
% >>> Particle = Body(differential=init_conditions, parent=Attractor)
% >>> geodesic = Geodesic(body=Particle, end_lambda=0.00205, step_size=5e-8)

% \end{lstlisting}
% We can see that for each timestep, we obtain eight values : $t$, $r$, $\theta$, $\phi$, $v_t$, $v_r$, $v_\theta$ and $v_\phi$ corrosponding to coordinate time, radial distance from the centre, polar angle, azimuth angle and their respective velocities w.r.t proper time $\tau$.

% Further, we plot the trajectory in 2D for vizualization purposes.
% \begin{lstlisting}[language=Python, caption=Plotting the trajectory]
% >>> obj = GeodesicPlotter()
% >>> obj.plot(geodesic)
% >>> obj.show()
% \end{lstlisting}

% \begin{figure}[h]
% 	\centering
% 	\includegraphics[scale=0.7]{figures/perihelion_cropped.png}
% 	%\centerline{\includegraphics{ftcbuild2.pdf}}
% 	\caption{Plot obtained from the above code showing the phenomenon of perihelion advance}
% 	\label{fig:schwarzschild}
% \end{figure}

% We observe the phenomenon of perihelion advance of a particle around a heavy body due to spacetime curavture. The orbit would have been purely eliptical if plotted in accordance with Newtonian Mechanics, however, the curvature of space-time around the heavy mass shifts the orbit with each revolution. In practicality, this phenomenon is not so pronounced as no particle approaches a heavy mass this closely. For instance, Mercury advances its perihelion at a rate of $42.98\pm 0.04$ arcsec/century.

@@

@@

% \begin{lstlisting}[language=Python, caption=Defining the Kerr Body and the particle and calculating geodesics]
% >>> a = 0.3 * u.m
% >>> attractor = Body(name="BH", mass=1.989e30 * u.kg, a=a)
% >>> coordinates = BoyerLindquistDifferential(49.95e5 * u.km, np.pi / 2 * u.rad, np.pi * u.rad, a=a)
% >>> particle = Body(differential=init_conditions, parent=attractor)
% >>> geodesic = Geodesic(body=particle, end_lambda=33932.90, step_size=1.2, metric=Kerr)
% \end{lstlisting}
% As evident from the code, API across different classes are made to be as homogeneous as possible.

% As a rotating black hole drags the space-time around it, a freely falling object acquires a angular velocity around the coordinate centre, i.e centre of the black hole. This effect is evident from the plot obtained from the above geodesics, in the same way as obtained one in the Schwarzschild example.

% \begin{figure}[h]
% 	\centering
% 	\includegraphics[scale=0.1]{figures/blank.jpg} %replace this when the actual plot becomes available
	%\centerline{\includegraphics{ftcbuild2.pdf}}
% 	\caption{Plot obtained from the above geodesics showing the phenomenon of frame dragging}
% 	\label{fig:kerr}
% \end{figure}

\subsubsection{Kerr-Newman Metric}\label{subsubsec:kerrnewman}
Kerr-Newman metric presents the most general case within single-body vacuum solutions of Einstein Field Equation. It generalizes all the quantities used to describe a black hole, as affirmed by "no hair" theorem, i.e. Charge $Q$, Angular Momentum $J$ and Mass $M$. The metric equation is given by,

@@

\subsection{COORDINATES}
The coordinates subpackage provides support for representing and transforming coordinates. Cartesian, Spherical and BoyerLindquist coordinate system can be accessed while using einsteinpy.

\subsection{SYMBOLIC}
General Relativity required heavy mathematical formulation to derive and describe various tensors attributing to various features essential for cogent description of an arbitrary space-time. We would describe the various quantities currently supported by this module while maintaining brevity. Each quantity has its own class, even scalars, which are treated as 0\textsuperscript{th} order tensors.

@@

% The defined class functions are :
% \begin{itemize}
% \item \texttt{change\_config} : returns a different instance of the same class with different index (contravariant or covariant) configuration. 
% \item \texttt{inv} : returns inverse of the metric
% \end{itemize}

Also, the library maintains an ever expanding predefined metrics for direct use in research purposes.

%\begin{lstlisting}[language=Python, caption=Importing a predefined metric]
%>>> from einsteinpy.symbolic.predefined import CMetric
%>>> metric = CMetric
% >>> metric.tensor()
% \end{lstlisting}

% \begin{center}
% $\displaystyle \left[\begin{matrix}- \frac{h{\left(y \right)}}{\left(x + y\right)^{2}} & 0 & 0 & 0\\0 & \frac{1}{\left(x + y\right)^{2} f{\left(x \right)}} & 0 & 0\\0 & 0 & \frac{1}{\left(x + y\right)^{2} h{\left(y \right)}} & 0\\0 & 0 & 0 & \frac{f{\left(x \right)}}{\left(x + y\right)^{2}}\end{matrix}\right]$
% \end{center}

% \begin{lstlisting}[language=Python, caption=Displaying the contravarint configuration of metric]
% >>> metric.change_config("uu").tensor()
% \end{lstlisting}

% \begin{center}
%  $\displaystyle \left[\begin{matrix}- \frac{\left(x + y\right)^{2}}{h{\left(y \right)}} & 0 & 0 & 0\\0 & \left(x + y\right)^{2} f{\left(x \right)} & 0 & 0\\0 & 0 & \left(x + y\right)^{2} h{\left(y \right)} & 0\\0 & 0 & 0 & \frac{\left(x + y\right)^{2}}{f{\left(x \right)}}\end{matrix}\right]$ \\
% \end{center}

@@

% Instance of the class can be made by directly passing a 3\textsuperscript{rd} order tensor or can be obtained using the following classmethods,
% \begin{itemize}
% \item \texttt{from\_metric} : Calculates christoffel symbols from a given instance of \hyperref[subsubsec:metric]{\textit{MetricTensor}} class.
% \end{itemize}

% The defined class function(s) are : 
% \begin{itemize}
% \item \texttt{change\_config} : returns a different instance of the same class with different index (contravariant or covariant) configuration. 
% \end{itemize}

@@

% Instance of the class can be made by directly passing a 4\textsuperscript{th} order tensor or can be obtained using the following classmethods
% \begin{itemize}
% \item \texttt{from\_metric} : Calculates Riemann Tensor from a given instance of \hyperref[subsubsec:metric]{\textit{MetricTensor}} class.
% \item \texttt{from\_christoffels} : Calculates Riemann Tensor from a given instance of \hyperref[subsubsec:chl]{\textit{ChristoffelSymbols}} class.
% \end{itemize}

% The defined class function(s) are : 
% \begin{itemize}
% \item \texttt{change\_config} : returns a different instance of the same class with different index (contravariant or covariant) configuration. 
% \end{itemize}

@@

% Instance of the class can be made by directly passing a 2\textsuperscript{nd} order tensor or can be obtained using the following classmethods
% \begin{itemize}
% \item \texttt{from\_metric} : Calculates Ricci Tensor from a given instance of \hyperref[subsubsec:metric]{\textit{MetricTensor}} class.
% \item \texttt{from\_christoffels} : Calculates Ricci Tensor from a given instance of \hyperref[subsubsec:chl]{\textit{ChristoffelSymbols}} class.
% \item \texttt{from\_riemann} : Calculates Ricci Tensor from a given instance of \hyperref[subsubsec:riemann]{\textit{RiemannCurvatureTensor}} class.
% \end{itemize}

% The defined class function(s) are : 
% \begin{itemize}
% \item \texttt{change\_config} : returns a different instance of the same class with different index (contravariant or covariant) configuration. 
% \end{itemize}

@@

% Instance of the class can be made by directly passing a symbolic expression or can be obtained using the following classmethods
% \begin{itemize}
% \item \texttt{from\_metric} : Calculates Ricci Scalar from a given instance of \hyperref[subsubsec:metric]{\textit{MetricTensor}} class.
% \item \texttt{from\_christoffels} : Calculates Ricci Scalar from a given instance of \hyperref[subsubsec:chl]{\textit{ChristoffelSymbols}} class.
% \item \texttt{from\_riemann} : Calculates Ricci Scalar from a given instance of \hyperref[subsubsec:riemann]{\textit{RiemannCurvatureTensor}} class.
% \item \texttt{from\_riccitensor} : Calculates Ricci Scalar from a given instance of \hyperref[subsubsec:riccitensor]{\textit{RicciTensor}} class.
% \end{itemize}

% The defined class properties are : 
% \begin{itemize}
% \item \texttt{expr} : returns the scalar symbolic expression contained in the class. 
% \end{itemize}

@@

\subsection{HYPERSURFACE}
This module provides basic computational functions which are essential for modelling the space-like hypersurface for any space-time geometry. Currently module has implementation for Schwarzschild geometry with conversion functions from Schwarzschild coordinates to 3-D spherical coordinates and some plotting utilities. The whole module is expected to extend for providing a super-class for implementations related to gravitational lensing simulations and numerical calculations of null geodesics on hypersurfaces and deviation angles for systems which are relevant to relativistic astrophysical problems. A small code snippet for plotting the hypersurface is as follows:
\begin{lstlisting}[language=Python, caption=Schwarzschild Embedding]
>>> surface_obj = SchwarzschildEmbedding(5.927e23 * u.kg)
>>> surface = HypersurfacePlotter(embedding=surface_obj, plot_type='surface')
>>> surface.plot()
>>> surface.show()
\end{lstlisting}

\subsection{PLOTTING}
The plotting ecosystem in EinsteinPy has two back-ends for two type of users. A static, matplotlib based plotter for publication quality plots and an interactive plotly based plotter for analysis and interpretation. \\
The plotting module is specifically designed keeping in mind the variety of IDEs and tools people working with Python use, for example python scripts, jupyter notebooks, spyder, pycharm etc. As plotly plots need JavasScript to render plots, it becomes difficult for the first time user to decide which backend to use. \texttt{StaticGeodesicPlotter} and \texttt{InteractiveGeodesicPlotter} undergo an internal switching depending on the environment variables and the appropriate environment aware plot is shown to the user for a better user experience and interface.\\
The plotting module throughout the applications follow a very consistent API. \begin{lstlisting}[language=Python, caption=Plotting Module API]
>>> data = Plotter()
>>> data.plot()
>>> data.show()
\end{lstlisting}

@@

\subsection{ Documentation and gallery} \label{subsec:doc}
The EinsteinPy Project strives to provide high quality, up-to-date and accurate documentation of the software. The documentation is highly approachable with easy and intuitive user interface. The documentation is self-updating with every change in the codebase, and is rendered automatically from the docstrings written for each function and classes. Along with the main API, new example codes for different functionalities is added to the repository frequently in the form of jupyter notebooks. Many sections like User Guide and Developer Guide has been written for easy on-boarding of new users and contributors.

@@

We thank all the people that have made EinsteinPy what it is today.  
The organisations ESA, Poliastro and Indian institute of Technology, Mandi provided us with significant support and have our deepest thanks.
We would specially like to thank Dr.Abhay Ashtekar and  Dr. Timothy A. Gonsalves for their constant motivation.
Lastly, thanks to all our contributors who contrubuted towards building EinsteinPy.

@@

@@

% \appendix
% \section{Appendix information}
